White Dwarf Mergers on Adaptive Meshes\\ II. Collisions

### 2017-09-09 papers/wdmerger_II/paper.tex
Fix the comment about whether we use retries
Update the description of the ODE integrators

In our \microphysics\ repository we provide several software options for
solving a set of coupled stiff ODEs. For this work we use an implementation of
the well known variable-order Richardson extrapolation method presented by \citet{stoer:1980},
that is similar to the integrator which ships with
the original versions of the networks mentioned above. Previous work using the \amrex\
codes typically used the classic \vode\ integrator \citep{vode}. We do maintain a version
of the software that is compatible with our software interfaces in the \microphysics\
repository, but we have largely shifted to integrators which are written in modern
Fortran as a consequence of our efforts to run our codes on GPUs. The Stoer and Bulirsch
integrator we provide satisfies this criterion; we also provide a rewrite of the VODE
BDF algorithm that uses modern Fortran features. The Stoer and Bulirsch algorithm
relies on a uniform relative error tolerance for all of the ODEs in the system, which
we set at $10^{-6}$ for all simulations described here.

@@

Now we return to a point we hinted at above. The timestep
will only actually satisfy the energy criterion
$\Delta t \leq f_e \tau_e$ and species criterion
$\Delta t \leq f_{X_k} \tau_{X_k}$ when the estimates for
$\dot{e}$ and $\dot{X_k}$ we generate are at least as large
as the actual rate of change of energy and mass fractions
over the timestep. However, this can assumption can fail
during periods of runaway burning when the rate of change
of these quantities is highly nonlinear. We may not want
to neglect the errors caused by this approximation
because they may build up over an extended period of nonlinear
evolution and perhaps substantially change the final results.
To this end, we have implemented a timestep retry option in
\castro, which re-computes an advance if it violated the
stability criteria as judged from the end of the timestep.
However we have found that the benefits for this problem are
small and thus we do not use it for the simulations here.

@@

>>>>

In our \microphysics\ repository we provide several software options for
solving a set of coupled stiff ODEs. For this work our primary integrator
is the classic \vode\ integrator \citep{vode}, of which we have a version
compatible with our software interfaces in the \microphysics\ repository.
We also employ an implementation of the well known variable-order Richardson
extrapolation method presented by \citet{stoer:1980}, that is similar to the
integrator which ships with the original versions of the networks mentioned
above. Our experience shows that relying on a single ODE integrator for the
variety of stiff ODEs that are present in a collision is insufficient, as we
sometimes run into integration failures. A combination of the two integrators
is usually sufficient, however: we first try \vode, and if that fails, we then
try the Stoer and Bulirsch method. Since it is possible for neither method to
succeed, we have one more backup scheme. Our specified error tolerances are
$10^{-6}$ for the energy and temperature ODEs, and $10^{-8}$ for the
nuclear abundance ODEs (we discuss the sensitivity to this choice in
\autoref{sec:parameters:tolerance}). We can loosen these tolerances in an
attempt to give the integrator an easier time. If both \vode\ and the Stoer
and Bulirsch algorithm fail, then we loosen the tolerances by a multiplicative
factor of 25\%, and then repeat the attempts with \vode\ and Stoer and Bulirsch.
We keep doing this until either we have loosened the tolerances enough for one
of the two methods to succeed, or until we have loosened the tolerances by two
orders of magnitude, at which point we finally give up. We have found that this
scheme is resilient enough to handle all of the simulations shown below.

@@

Now we return to a point we hinted at above. The timestep
will only actually satisfy the energy criterion
$\Delta t \leq f_e \tau_e$ and species criterion
$\Delta t \leq f_{X_k} \tau_{X_k}$ when the estimates for
$\dot{e}$ and $\dot{X_k}$ we generate are at least as large
as the actual rate of change of energy and mass fractions
over the timestep. However, this can assumption can fail
during periods of runaway burning when the rate of change
of these quantities is highly nonlinear. We may not want
to neglect the errors caused by this approximation
because they may build up over an extended period of nonlinear
evolution and perhaps substantially change the final results.
To this end, we have implemented a timestep retry option in
\castro, which re-computes an advance if it violated the
stability criteria as judged from the end of the timestep.
However we have found that the benefits for this problem are
small.

@@

\subsection{Integration Error Tolerance}
\label{sec:parameters:tolerance}

%TODO: add discussion of how ODE tolerances affect the answer
